---
title: GitHub Metrics
output: md_document
---

```{r include = FALSE}
library(gh)
library(tidyverse)

knitr::opts_chunk$set(
  echo = FALSE
  , message = FALSE
  , warning = FALSE
  , error = FALSE
)
```

```{r}
account_established <- as.Date("2019-08-01")
```

# Overview

This file was updated on `r Sys.Date()`.

This document will provide a basic overview of the volume of and interaction with the CAS GitHub organizational account. The account was established in August of 2019.

## Repos

```{r}
lst_cas_repos <- gh("GET /orgs/{org}/repos", org = "CASact", .limit = Inf)

repo_cols <- c(
  'name'
  , 'description'
  , 'fork'
  , 'created_at'
  , 'updated_at'
  , 'pushed_at'
  , 'stargazers_count'
  , 'watchers'
  , 'watchers_count'
  , 'language'
  , 'forks'
  , 'forks_count'
  , 'open_issues'
  , 'open_issues_count'
)

tbl_repo <- tibble(repo = lst_cas_repos) %>% 
  unnest_wider(repo) %>% 
  select_at(repo_cols) %>% 
  mutate(
    created_at = as.Date(created_at)
  )
```


```{r}
repo_count <- nrow(tbl_repo)
```

As a reminder, a repository --- typically referred to as a "repo", for short --- is the unit of measure for a single project. A repo may contain more than one file, and in fact, most repos contain multiple files.

## By date

As of `r Sys.Date()`, there are `r repo_count` repositories on the site.

The cumulative count of repos is shown in Figure \@ref(fig:cumulative_repo). We note that we begin with ten repos as of the day that the account was established. This is a result of several repos which predate the CAS' account having been transferred from other owners to the CAS.

```{r cumulative_repo}
tbl_repo %>% 
  arrange(created_at) %>% 
  mutate(
    num_repos = nrow(.) %>% 
      seq_len() 
  ) %>%
  filter(created_at > account_established) %>% 
  group_by(created_at) %>% 
  summarise(
    num_repos = max(num_repos)
  ) %>% 
  bind_rows(
    tibble(created_at = Sys.Date(), num_repos = repo_count)
  ) %>% 
  ggplot(aes(created_at, num_repos)) + 
  geom_step() + 
  scale_y_continuous(limits = c(0, NA), breaks = seq(0, 20, by = 5), labels = scales::label_comma(accuracy = 1))
```

## By language

Where there is a predominant language for the repo, GitHub will indicate it. Of those repos with a single predominant language, the number of repos is as follows.

```{r}
tbl_repo %>% 
  filter(!is.na(language)) %>% 
  mutate(
    language = language %>% 
      fct_infreq() %>% 
      fct_rev()
  ) %>% 
  ggplot(aes(language)) + 
  geom_bar() +
  coord_flip()
```

## Watchers

```{r}
tbl_repo %>% 
  filter(watchers > 0) %>% 
  mutate(name = fct_reorder(name, watchers)) %>% 
  ggplot(aes(name, watchers)) + 
  geom_bar(stat = 'identity') +
  coord_flip()
```

<!--
## Stars

```{r eval = FALSE}
tbl_repo %>% 
  filter(stargazers_count > 0) %>% 
  mutate(name = fct_reorder(name, stargazers_count)) %>% 
  ggplot(aes(name, stargazers_count)) + 
  geom_bar(stat = 'identity') +
  coord_flip()
```
-->

## Forks

```{r}
tbl_repo %>% 
  filter(forks > 0) %>% 
  mutate(name = name %>% as.factor() %>% fct_reorder(forks)) %>% 
  ggplot(aes(name, forks)) + 
  geom_bar(stat = 'identity') +
  coord_flip()
```

<!-- Maybe add some stuff about words in the description or some shit. -->

# Contributions

## Commits

```{r}
get_repo_commits <- function(repo) {
  
  lst_commits <- gh(
    "GET /repos/{owner}/{repo}/commits"
    , owner = "casact"
    , repo = repo
    , affiliation = "all"
    , .limit = Inf)
  
  tbl_repo_commits <- lst_commits %>% 
    map_dfr(function(x) {
      tibble(
        repo = repo
        , sha = x$sha
        , author = x$author$login %>% ifelse(is.null(.), NA, .)
        # , datetime = x$commit$author$date %>% as.Date
        , date = x$commit$author$date %>% as.Date()
      )
    })
}

tbl_commits <- tbl_repo$name %>% 
  map_dfr(get_repo_commits)
```

There have been `r nrow(tbl_commits) %>% format(big.mark = ",")` commits in total.

```{r }
tbl_commits_by_date <- tbl_commits %>% 
  count(repo, date, name = 'commits')

tbl_commits_top_ten <- tbl_commits_by_date %>% 
  group_by(repo) %>% 
  summarise(total_commits = sum(commits)) %>% 
  arrange(desc(total_commits)) %>% 
  mutate(
    rank = seq_len(nrow(.))
  ) %>% 
  head(10) %>% 
  ungroup()

tbl_commits_by_date <- left_join(
  tbl_commits_by_date, tbl_commits_top_ten, by = "repo"
)
```

The top ten repos by number of commits are:

```{r}
tbl_commits_top_ten %>% 
  select(-rank) %>% 
  knitr::kable()
```

The cumulative growth of commits over time has been:

```{r}
tbl_commits_by_date %>% 
  group_by(date) %>% 
  summarise(total_commits = sum(commits)) %>% 
  arrange(date) %>% 
  mutate(cumulative_commits = cumsum(total_commits)) %>% 
  filter(date > account_established) %>% 
  ggplot(aes(date, cumulative_commits)) + 
  # geom_point() + 
  geom_line() +
  NULL
```

<!-- Show commits over time for each repo separately. 

```{r}
tbl_commits_by_date %>% 
  filter(date > account_established) %>% 
  ggplot(aes(date, commits)) + 
  geom_point()
```

```{r}
tbl_commits_by_date %>% 
  filter(rank < 10) %>% 
  filter(date > account_established) %>% 
  ggplot(aes(date, commits, color = repo)) + 
  geom_point()
```


```{r}
tbl_commits_by_date %>% 
  filter(rank < 10) %>% 
  group_by(repo) %>% 
  arrange(date) %>% 
  mutate(cumulative_commits = cumsum(commits)) %>% 
  filter(date > account_established) %>% 
  ggplot(aes(date, cumulative_commits, color = repo)) + 
  geom_point() + 
  geom_line() +
  NULL
  facet_wrap(~repo, scales = 'free_y')
```

```{r eval = FALSE}
lst_commit_activity <- gh(
  "GET /repos/{owner}/{repo}/stats/commit_activity"
  , owner = "casact"
  , repo = "imaginator"
  , .limit = Inf
)
```
-->

## Commiters

The ten most frequent contributors to various repos are:

```{r}
tbl_commits %>% 
  filter(!is.na(author)) %>% 
  group_by(author) %>% 
  count() %>% 
  arrange(desc(n)) %>% 
  head(10) %>% 
  knitr::kable()
```

Authors who have worked on multiple repositories are:

```{r}
tbl_commits %>% 
  filter(!is.na(author)) %>% 
  select(author, repo) %>% 
  unique() %>% 
  group_by(author) %>% 
  count() %>% 
  arrange(desc(n)) %>% 
  filter(n > 1) %>% 
  head(10) %>% 
  knitr::kable()
```

## Pull requests

```{r}
lst_pulls <- gh(
  "GET /repos/{owner}/{repo}/pulls"
  , owner = "casact"
  , repo = "imaginator"
  , state = "all"
  , .limit = Inf)
```

Pull requests take place when a GitHub user submits an improvement or correction to be considered by the maintainer of the repo.

Metrics coming soon.
